Scalar and Representative Observables for Polarimetric SAR Data

### 2013-11-05 Remote_Sensing_Letter/paper.tex
RSL template applied => too long!

%ifdef FINAL
\documentclass[journal]{IEEEtran}
%endif
%ifdef REVIEW
\documentclass[journal,12pt,draftcls,onecolumn]{IEEEtran}
%endif
%\documentclass[a4paper, 10pt, conference]{ieeeconf}

@@

\newsavebox\IBox
\savebox\IBox{\includegraphics[width=1.5in,height=1.5in]{../images/verify_change_ratio_model_on_RADARSAT2_3d.eps}}

@@

\newcommand{\plotWithLegend}[2]{          
          \begin{pspicture}[showgrid=false](\column,\row)% set showgrid=false for the final
	    \rput[bl](0,0){\includegraphics[width=1.5in,height=1.5in]{#1}}% \usebox\IBoxCRThreeR
	    %\rput(5,1.5){\footnotesize{#2}}
	    %\rput{90}(1.5,5){\footnotesize{pdf / histogram}}
	    \psline[linecolor=plot](5,8)(6,8)
	    \psline[linestyle=dashed](5,9)(6,9)%another line style: dotted
	    \rput(8,8){\scriptsize{model}}
	    \rput(8,9){\scriptsize{data}}            
          \end{pspicture}
}

@@

\usepackage[center]{caption}%for captions
\usepackage[caption=false,font=scriptsize]{subfig} %for subfigures

@@

%\author{Thanh-Hai Le, Ian McLoughlin}
\author{Thanh-Hai~Le,
        Ian~McLoughlin, 
	and Chan-Hua~Vun%
\thanks{Thanh-Hai~Le and Chan-Hua~Vun are with School of Computer Engineering, 
Nanyang Technological University, Singapore. Ian~McLoughlin is with School of Information Science and Technology,
University of Science and Technology of China.
}% <-this % stops a space
%\thanks{The authors wish to thank Dr. Ken-Yoong Lee and Dr. Timo Brestchneider of EADS Innovation-Works Singapore for 
%	providing us the RADAR-SAT2 imagery used in this paper. }% <-this % stops a space
\thanks{Manuscript received ?, 2013; revised ?.}}

\markboth{IEEE Journal on Selected Topic in Applied Earth Observation and Remote Sensing,~Vol.~?, No.~?, ?~2013}%
{ Le \MakeLowercase{\textit{et al.}}:  Scalar and Representative Observables for Polarimetric SAR Data}

@@

@@

\begin{IEEEkeywords}
Polarimetric Synthetic Aperture Radar, Electromagnetic Modeling, Multidimensional Signal Processing  
\end{IEEEkeywords}

\IEEEpeerreviewmaketitle

@@

%How do others solve this problem?
%The complex POLSAR data have been statistically modelled as following the complex Wishart distribution,
%  which apparently is multidimensional, complex and thus not very intuitive.
There has been a few scalar observables with accompanying statistical models proposal for POLSAR \cite{Conradsen_2003_TGRS_4, Alberga_2008_IJRS_4129, Joughin_1994_TGRS_562, Lee_1994_TGRS_1017, Touzi_1996_TGRS_519, Lopez-Martinez_2003_TGRS_2232, Erten_2012_Sensors_2766},
  but none of them is able to provide meaningful scalar discrimination measures.
%  but none of them has led to scalar discrimination measures proposal.
As such no observable has been widely accepted as being highly representative of this multi-dimensional data,
  which severely limits their applicability in practical data processing applications.
From an alternative approach, a few POLSAR discrimination measures have been proposed,
  but all are based on the likelihood ratio statistics.
This statistical test should be based on an exact and consistent statistical distribution
  but so far has only shown to be based on an asymptotic distribution.

@@

%This section outlines related publication.
Section \ref{sec:lit_models} reviews various statistical models used for different scalar observables for POLSAR.
  where it is shown that none are able to provide statistically consistent discrimination measures. 
Section \ref{sec:lit_measures} further strengthens those findings by discussing discrimination measures that have been proposed for POLSAR,
  and demonstrates that almost all of them are based on the likelihood statistical test for complex Wishart distribution.
While an exact statistical distribution is expected to be necessary for a test,
  only an asymptotic distribution is used in the underlying approach \cite{Conradsen_2003_TGRS_4}.

@@

Different target decomposition theorems have identified many possible scalar observables for complex POLSAR data.
In \cite{Alberga_2008_IJRS_4129}, the performance of different scalar POLSAR observables is evaluated for classification purposes.
While many scalar observables for POLSAR were presented, their corresponding statistical models and classifiers were not available.
Furthermore, at its conclusion, the paper indicated that it is impossible to identify a single best representation.
Although, to be fair, the observables were identified for describing a decomposed portion of the complex POLSAR data,
  rather than providing a unified representation of the POLSAR data.

Using a different approach, given that the joint distribution for POLSAR is known to be the multi-variate complex Wishart distribution,
  it is possible to derive the scalar statistical models for some univariate POLSAR observables.
%This becomes an alternative approach used in the study of POLSAR data.
However, such derivations are not trivial tasks, and so far, only a handful of such statistical models have been proposed, including the following:
  \begin{enumerate}
  \item cross-pol ratio $r_{HV/HH} = |S_{HV}|^2/|S_{HH}|^2$ \cite{Joughin_1994_TGRS_562},
  \item co-pol ratio $r_{VV/HH} = |S_{VV}|^2/|S_{HH}|^2$ \cite{Joughin_1994_TGRS_562},
  \item co-pol phase difference $\phi_{VV/HH} = arg(S_{VV}S_{HH}^*) $ \cite{Joughin_1994_TGRS_562} \cite{Lee_1994_TGRS_1017},
  \item magnitude $g=|avg(S_{pq}S_{rs}^*)|$ \cite{Lee_1994_TGRS_1017},
  \item normalized magnitude $\xi = \frac{|avg(S_{pq}S_{rs}^*)|}{\sqrt{avg(|S_{pq}|^2) avg(|S_{rs}|^2)}}$ \cite{Lee_1994_TGRS_1017},
  \item intensity ratio $w = avg(|S_{pq}|^2)/avg(|S_{rs}|^2)$ \cite{Lee_1994_TGRS_1017},
  \item and the Stokes parameters $S_i,0 \leq i \leq 3$ \cite{Touzi_1996_TGRS_519}. 
  \end{enumerate}
More recently, statistical models for
  each element of the POLSAR covariance matrix, i.e. $S_{pq}S_{rs}^*$, \cite{Lopez-Martinez_2003_TGRS_2232}
  as well as for the largest eigen-value of the covariance matrix $\lambda_1$ \cite{Erten_2012_Sensors_2766} have been proposed.

@@

In the field of POLSAR, the Wishart distance is probably the most widely used, as part of the well-known Wishart classifier \cite{Lee_1999_TGRS}.
It is defined \cite{Lee_1994_IJRS_2299} as:
\begin{equation}
  d(C_x,C_y) = \ln|C_y| + tr(C_xC_y^{-1})
\end{equation},
where $tr(C)$ denotes the trace of the matrix C. 
As a measure of distance, its main disadvantage is that $d(C_y,C_y) = \ln|C_y| \neq 0$.

Recent works have suggested alternative dissimilarity measures including the symmetric and asymmetric refined Wishart distance \cite{Anfinsen_2007_ESA_POLINSAR},
\begin{align}
  d(C_x,C_y) &= \frac{1}{2} tr(C_x^{-1}C_y + C_y^{-1}C_x) - d \\
    d(C_x,C_y) &= \ln|C_x| - \ln|C_y| + tr(C_xC_y^{-1}) - d
\end{align}
the Bartlett distance \cite{Kersten_2005_TGRS_519},
  \begin{align}
  d(C_x,C_y) &= 2 \ln |C_{x+y}| - \ln |C_x| - \ln |C_y| - 2d\ln2
  \end{align}
the Bhattacharyya distance \cite{Lee_2011_IGARSS_3740},
\begin{equation}
  r(C_x,C_y) = \frac{|C_x|^{1/2} |C_y|^{1/2}}{|(C_x+C_y)/2|}
\end{equation}
and the Wishart Statistical test distance \cite{Cao_2007_TGRS_3454}
\begin{equation}
  d(C_x,C_y) = (L_x + L_y) \ln|C| - L_x \ln|C_x| - L_y\ln|C_y|
\end{equation}

Closer examination of these dissimilarity measures reveals that most of them are related in some ways. %to each other.
The Bhattacharyya distance is easily shown to be related to the Barlett distance.
At the same time the Barlett distance can be considered a special case of the Wishart Statistical Test distance,
  when the two data sets have the same number of looks, i.e. $L_x=L_y$.
The close relation among the measures is further supported by the fact that
  all of their publications refered the same statistical model developed in \cite{Conradsen_2003_TGRS_4} as the foundation.
In \cite{Conradsen_2003_TGRS_4}, to determine if the two scaled multi-look POLSAR covariance matrixes $Z_x$ and $Z_y$,
  which have $L_x$ and $L_y$ as the corresponding number of looks,
  come from the same underlying stochastic process,
the likelihood ratio statistics for POLSAR covariance matrix is considered:  
\begin{equation*}
  Q = \frac{(L_x+L_y)^{d \cdot (L_x+L_y)}}{L_x^{d \cdot L_x} L_y^{d \cdot L_y}} \frac{|Z_x|^{L_x} |Z_y|^{L_y} }{|Z_x+Z_y|^{(L_x+L_y)}}
\end{equation*}

@@

To detect changes, a test statistic is developed for this discrimination measure.
This means a distribution is to be derived for the dissimilarity measure.
However, in the original work \cite{Conradsen_2003_TGRS_4}, only an asymptotic distribution was used.
This paper proposes a statistical model for the determinant of the POLSAR covariance matrix $|C_v|$
  which is capable of providing an exact distribution for the test.

@@

The approach taken in this paper differs by applying the homoskedastic log transformation  on a less-than-well-known result for the determinant of the covariance matrix.
Goodman \cite{Goodman_1963_AMS_178} proved
that the ratio between the observable and expected values of the sample covariance matrix determinants
  behaves like a product of $d$ chi-squared random variables with different degrees of freedom: 
\begin{equation}
\chi^d_L = (2L)^d \frac{|C_v|}{|\Sigma_v|} \sim \prod_{i=0}^{d-1} \chi (2L-2i)
\label{eqn:prod_chi_squared_rv}  
\end{equation}

@@

Moreover, since the average and variance of these chi-squared distributions %and log-chi-squared distribution
  are known to be constant, i.e. $avg \left[ \chi(2L) \right] = 2L$ and $var \left[ \chi(2L) \right] = 4L$,
  the product and summation of these random variables also have fixed summary statistics.
Specifically:
{\footnotesize
\begin{align*}
  avg \left[ \prod^{d-1}_{i=0} \chi(2L-2i) \right] &= 2^d \cdot \prod^{d-1}_{i=0} (L-i), \\
  var \left[ \prod^{d-1}_{i=0} \chi(2L-2i) \right] &= \prod^{d-1}_{i=0} 4(L-i)(L-i+1) - \prod^{d-1}_{i=0} 4(L-i)^2, %\\
%  avg \left[ \sum^{d-1}_{i=0} \Lambda(2L-2i) \right] &= d \cdot \ln{2} + \sum^{d-1}_{i=0} \psi^0(L-i), \\
%  var \left[ \sum^{d-1}_{i=0} \Lambda(2L-2i) \right] &= \sum^{d-1}_{i=0} \psi^1(L-i)
\end{align*}
}

Combining these results with Eqn. \ref{eqn:determinant_distribution} %and \ref{eqn:log_determinant_distribution}
  , we have:
{\footnotesize
\begin{align}
  avg \left[ |C_v| \right]  &= \frac{|\Sigma_v|}{L^d} \prod^{d-1}_{i=0} (L-i)\\
  var \left[ |C_v| \right]  &=   \frac{|\Sigma_v|^2 \left[ \prod^{d-1}_{i=0} (L-i)(L-i+1) - \prod^{d-1}_{i=0} (L-i)^2 \right] }{L^{2d}} \label{eqn:var_det_is_heteroskedastic}%\\
%  avg \left[ \ln |C_v| \right] &= \ln |\Sigma_v| - d \cdot \ln{L}  + \sum^{d-1}_{i=0} \psi^0(L-i) \label{eqn:avg_log_det} \\
%  var \left[ \ln |C_v| \right] &=  \sum^{d-1}_{i=0} \psi^1(L-i) \label{eqn:var_log_det_is_homoskedastic}
\end{align}
}

@@

Similar to the way intensity-ratio is proposed as the discrimination measure for the multiplicative and heteroskedastic SAR intensity \cite{Rignot_1993_TGRS_896},
in this paper, the determinant-ratio and the change-ratio are also proposed as a discrimination measure for the POLSAR data.
  which is shown above to also suffer from the multiplicative and heteroskedastic phenomena.
%Similar to the way dispersion and contrast is defined for one-dimensional SAR \cite{Le_2010_ACRS},
%  this section introduces the consistent sense of distance for POLSAR.

@@

Statistical models have always been the foundation for deriving discrimination measures.
This is true for both fields of SAR and POLSAR.
For SAR, where the field is more or less settled on the issue,
  the statistical model for SAR intensity has been used to derive the most widely used intensity-ratio discrimination measure \cite{Rignot_1993_TGRS_896}.
For POLSAR, where the research field is much less mature,
  the same case should apply, 
  except that so far only asymptotic distributions have been derived for the most common foundation, i.e. the likelihood test statistics \cite{Conradsen_2003_TGRS_4}.

With the insight gained from section \ref{sec:sar_special_case_of_polsar}, this section presents a few results.
%  it should be clear that the proposed model can also be used to derive discrimination measures for POLSAR.
First, similar to the way that the statistical model for SAR intensity has been used to provide the foundation for SAR discrimination measures, e.g. intensity-ratio \cite{Rignot_1993_TGRS_896},
  the statistical model proposed for POLSAR covariance matrix determinant is to be reviewed as providing the foundation for POLSAR discrimination measures, i.e. the likelihood test statistics.
And second, new discrimination measures for POLSAR may be derived by learning from the existing SAR discrimination measures.

As for the first matter, in view of the models given in Eqn \ref{eqn:determinant_distribution},
  the likelihood test statistics presented in \cite{Conradsen_2003_TGRS_4} and rewritten in Eqns \ref{eqn:ori_likelyhood_stats} \& \ref{eqn:log_likelyhood_stats}
can be expressed as:
\begin{align*}
  \ln{Q} &\sim  k + L_x \Lambda^d_{L_x} + L_y \Lambda^d_{L_y} - (L_x + L_y) \Lambda^d_{(L_x + L_y)} \\
  Q &\sim e^k \frac{(\chi^d_{L_x})^{L_x} \cdot (\chi^d_{L_y})^{L_y}}{(\chi^d_{L_x + L_y})^{L_x + L_y}}   
\end{align*}
where $k = d \left[ (L_x + L_y) \ln(L_x + L_y) - L_x \ln{L_x} - L_y \ln{L_y} \right]$.
This, in essence, derives an exact statistical distribution for the likelihood test statistics,
  as opposed to the asymptotic distribution derived in \cite{{Conradsen_2003_TGRS_4}}.

@@

This section aims to verify the theoretical models given in Eqns. \ref{eqn:determinant_distribution}, \ref{eqn:determinant_ratio_distribution} and \ref{eqn:change_ratio_distribution} against practical data.
All of these models require the estimation of two parameters from the captured data.
These are (i) the dimensional number $d$ and (ii) the look number $L$.
The dimensional number $d$ is related to the type of (POL)SAR data captured,
  with $d=1,2,3$ corresponding to the cases of SAR, partial and full polarimetric SAR data, respectively.
The look-number $L$ can be taken either as nominally stated by the data provider,
  or it can be estimated using the technique proposed by \cite{Anfinsen_2009_TGRS_3795}.

@@

Since it has been shown in Section \ref{sec:sar_special_case_of_polsar} that the model for the case of $d=1$ matches exactly with the traditional model for SAR intensity,
  the validation of the proposed $d=1$ model is quite straightforward. %only a simple validation is presented for this case.  
Fig. \ref{fig:verify_POLSAR_model_1D} presents the results of a test performed for the stated purpose.
In the test, the intensity of single-channel SAR data (HH) for sample homogeneous areas in both datasets is extracted.
The histograms for both the intensity and intensity ratio %the log-distance and contrast
  are then plotted against the theoretical PDF given above.
%Also presented are the plots of the same test applied on a homogeneous patch inside the nine-look processed RADARSAT2 image.
The plots are obtained having set ENL to an estimated value obtained using the technique presented in \cite{Anfinsen_2009_TGRS_3795}.
In all cases, the good visual match between the actual data and model distribution tends to validate the proposed model. %providing us a simple validation of the result.

>>>>

\documentclass[]{tRSL2e}

%\documentclass[journal]{IEEEtran}
%\documentclass[a4paper, 10pt, conference]{ieeeconf}

@@

\newsavebox\IBox
\savebox\IBox{\includegraphics[width=2.5in,height=2.5in]{../images/verify_change_ratio_model_on_RADARSAT2_3d.eps}}

@@

\newcommand{\plotWithLegend}[2]{          
          \begin{pspicture}[showgrid=false](\column,\row)% set showgrid=false for the final
	    \rput[bl](0,0){\includegraphics[width=2.5in,height=2.5in]{#1}}% \usebox\IBoxCRThreeR
	    %\rput(5,1.5){\footnotesize{#2}}
	    %\rput{90}(1.5,5){\footnotesize{pdf / histogram}}
	    \psline[linecolor=plot](5,8)(6,8)
	    \psline[linestyle=dashed](5,9)(6,9)%another line style: dotted
	    \rput(8,8){model}
	    \rput(8,9){data}            
          \end{pspicture}
}

@@

\usepackage[center]{caption}%for captions
\usepackage{subcaption}
%\usepackage[caption=false,font=scriptsize]{subfig} %for subfigures

@@

@@

\author{
  Thanh-Hai~Le$^{\ast}$$\dag$\thanks{$^\ast$Corresponding
author. Email: le\_thanh\_hai@hotmail.com \vspace{6pt}},
  and Ian~McLoughlin${\ddag}$, 
  and Chan-Hua~Vun${\dag}$\\
\vspace{6pt} $\dag$School of Computer Engineering, Nanyang Technological University, Singapore.\\
$\ddag$ School of Information Science and Technology,
University of Science and Technology of China.\\
\vspace{6pt}\received{November 2013} }

@@

@@

%How do others solve this problem?
%The complex POLSAR data have been statistically modelled as following the complex Wishart distribution,
%  which apparently is multidimensional, complex and thus not very intuitive.
There has been a few scalar observables with accompanying statistical models proposal for POLSAR \citep{Conradsen_2003_TGRS_4, Alberga_2008_IJRS_4129, Joughin_1994_TGRS_562, Lee_1994_TGRS_1017, Touzi_1996_TGRS_519, Lopez-Martinez_2003_TGRS_2232, Erten_2012_Sensors_2766},
  but none of them is able to provide meaningful scalar discrimination measures.
%  but none of them has led to scalar discrimination measures proposal.
As such no observable has been widely accepted as being highly representative of this multi-dimensional data,
  which severely limits their applicability in practical data processing applications.
From an alternative approach, a few POLSAR discrimination measures have been proposed,
  but all are based on the likelihood ratio statistics.
This statistical test should be based on an exact and consistent statistical distribution
  but so far has only shown to be based on an asymptotic distribution.

@@

%This section outlines related publication.
Section \ref{sec:lit_models} reviews various statistical models used for different scalar observables for POLSAR.
  where it is shown that none are able to provide statistically consistent discrimination measures. 
Section \ref{sec:lit_measures} further strengthens those findings by discussing discrimination measures that have been proposed for POLSAR,
  and demonstrates that almost all of them are based on the likelihood statistical test for complex Wishart distribution.
While an exact statistical distribution is expected to be necessary for a test,
  only an asymptotic distribution is used in the underlying approach \citep{Conradsen_2003_TGRS_4}.

@@

Different target decomposition theorems have identified many possible scalar observables for complex POLSAR data.
In \citep{Alberga_2008_IJRS_4129}, the performance of different scalar POLSAR observables is evaluated for classification purposes.
While many scalar observables for POLSAR were presented, their corresponding statistical models and classifiers were not available.
Furthermore, at its conclusion, the paper indicated that it is impossible to identify a single best representation.
Although, to be fair, the observables were identified for describing a decomposed portion of the complex POLSAR data,
  rather than providing a unified representation of the POLSAR data.

Using a different approach, given that the joint distribution for POLSAR is known to be the multi-variate complex Wishart distribution,
  it is possible to derive the scalar statistical models for some univariate POLSAR observables.
%This becomes an alternative approach used in the study of POLSAR data.
However, such derivations are not trivial tasks, and so far, only a handful of such statistical models have been proposed, including the following:
  \begin{enumerate}
  \item cross-pol ratio $r_{HV/HH} = |S_{HV}|^2/|S_{HH}|^2$ \citep{Joughin_1994_TGRS_562},
  \item co-pol ratio $r_{VV/HH} = |S_{VV}|^2/|S_{HH}|^2$ \citep{Joughin_1994_TGRS_562},
  \item co-pol phase difference $\phi_{VV/HH} = arg(S_{VV}S_{HH}^*) $ \citep{Joughin_1994_TGRS_562} \citep{Lee_1994_TGRS_1017},
  \item magnitude $g=|avg(S_{pq}S_{rs}^*)|$ \citep{Lee_1994_TGRS_1017},
  \item normalized magnitude $\xi = \frac{|avg(S_{pq}S_{rs}^*)|}{\sqrt{avg(|S_{pq}|^2) avg(|S_{rs}|^2)}}$ \citep{Lee_1994_TGRS_1017},
  \item intensity ratio $w = avg(|S_{pq}|^2)/avg(|S_{rs}|^2)$ \citep{Lee_1994_TGRS_1017},
  \item and the Stokes parameters $S_i,0 \leq i \leq 3$ \citep{Touzi_1996_TGRS_519}. 
  \end{enumerate}
More recently, statistical models for
  each element of the POLSAR covariance matrix, i.e. $S_{pq}S_{rs}^*$, \citep{Lopez-Martinez_2003_TGRS_2232}
  as well as for the largest eigen-value of the covariance matrix $\lambda_1$ \citep{Erten_2012_Sensors_2766} have been proposed.

@@

In the field of POLSAR, the Wishart distance is probably the most widely used, as part of the well-known Wishart classifier \citep{Lee_1999_TGRS}.
It is defined \citep{Lee_1994_IJRS_2299} as:
\begin{equation}
  d(C_x,C_y) = \ln|C_y| + tr(C_xC_y^{-1})
\end{equation},
where $tr(C)$ denotes the trace of the matrix C. 
As a measure of distance, its main disadvantage is that $d(C_y,C_y) = \ln|C_y| \neq 0$.

Recent works have suggested alternative dissimilarity measures including the symmetric and asymmetric refined Wishart distance \citep{Anfinsen_2007_ESA_POLINSAR},
\begin{align}
  d(C_x,C_y) &= \frac{1}{2} tr(C_x^{-1}C_y + C_y^{-1}C_x) - d \\
    d(C_x,C_y) &= \ln|C_x| - \ln|C_y| + tr(C_xC_y^{-1}) - d
\end{align}
the Bartlett distance \citep{Kersten_2005_TGRS_519},
  \begin{align}
  d(C_x,C_y) &= 2 \ln |C_{x+y}| - \ln |C_x| - \ln |C_y| - 2d\ln2
  \end{align}
the Bhattacharyya distance \citep{Lee_2011_IGARSS_3740},
\begin{equation}
  r(C_x,C_y) = \frac{|C_x|^{1/2} |C_y|^{1/2}}{|(C_x+C_y)/2|}
\end{equation}
and the Wishart Statistical test distance \citep{Cao_2007_TGRS_3454}
\begin{equation}
  d(C_x,C_y) = (L_x + L_y) \ln|C| - L_x \ln|C_x| - L_y\ln|C_y|
\end{equation}

Closer examination of these dissimilarity measures reveals that most of them are related in some ways. %to each other.
The Bhattacharyya distance is easily shown to be related to the Barlett distance.
At the same time the Barlett distance can be considered a special case of the Wishart Statistical Test distance,
  when the two data sets have the same number of looks, i.e. $L_x=L_y$.
The close relation among the measures is further supported by the fact that
  all of their publications refered the same statistical model developed in \citep{Conradsen_2003_TGRS_4} as the foundation.
In \citep{Conradsen_2003_TGRS_4}, to determine if the two scaled multi-look POLSAR covariance matrixes $Z_x$ and $Z_y$,
  which have $L_x$ and $L_y$ as the corresponding number of looks,
  come from the same underlying stochastic process,
the likelihood ratio statistics for POLSAR covariance matrix is considered:  
\begin{equation*}
  Q = \frac{(L_x+L_y)^{d \cdot (L_x+L_y)}}{L_x^{d \cdot L_x} L_y^{d \cdot L_y}} \frac{|Z_x|^{L_x} |Z_y|^{L_y} }{|Z_x+Z_y|^{(L_x+L_y)}}
\end{equation*}

@@

To detect changes, a test statistic is developed for this discrimination measure.
This means a distribution is to be derived for the dissimilarity measure.
However, in the original work \citep{Conradsen_2003_TGRS_4}, only an asymptotic distribution was used.
This paper proposes a statistical model for the determinant of the POLSAR covariance matrix $|C_v|$
  which is capable of providing an exact distribution for the test.

@@

The approach taken in this paper differs by applying the homoskedastic log transformation  on a less-than-well-known result for the determinant of the covariance matrix.
Goodman \citep{Goodman_1963_AMS_178} proved
that the ratio between the observable and expected values of the sample covariance matrix determinants
  behaves like a product of $d$ chi-squared random variables with different degrees of freedom: 
\begin{equation}
\chi^d_L = (2L)^d \frac{|C_v|}{|\Sigma_v|} \sim \prod_{i=0}^{d-1} \chi (2L-2i)
\label{eqn:prod_chi_squared_rv}  
\end{equation}

@@

Moreover, since the average and variance of these chi-squared distributions %and log-chi-squared distribution
  are known to be constant, i.e. $avg \left[ \chi(2L) \right] = 2L$ and $var \left[ \chi(2L) \right] = 4L$,
  the product and summation of these random variables also have fixed summary statistics.
Specifically:
{
\begin{align*}
  avg \left[ \prod^{d-1}_{i=0} \chi(2L-2i) \right] &= 2^d \cdot \prod^{d-1}_{i=0} (L-i), \\
  var \left[ \prod^{d-1}_{i=0} \chi(2L-2i) \right] &= \prod^{d-1}_{i=0} 4(L-i)(L-i+1) - \prod^{d-1}_{i=0} 4(L-i)^2, %\\
%  avg \left[ \sum^{d-1}_{i=0} \Lambda(2L-2i) \right] &= d \cdot \ln{2} + \sum^{d-1}_{i=0} \psi^0(L-i), \\
%  var \left[ \sum^{d-1}_{i=0} \Lambda(2L-2i) \right] &= \sum^{d-1}_{i=0} \psi^1(L-i)
\end{align*}
}

Combining these results with Eqn. \ref{eqn:determinant_distribution} %and \ref{eqn:log_determinant_distribution}
  , we have:
{
\begin{align}
  avg \left[ |C_v| \right]  &= \frac{|\Sigma_v|}{L^d} \prod^{d-1}_{i=0} (L-i)\\
  var \left[ |C_v| \right]  &=   \frac{|\Sigma_v|^2 \left[ \prod^{d-1}_{i=0} (L-i)(L-i+1) - \prod^{d-1}_{i=0} (L-i)^2 \right] }{L^{2d}} \label{eqn:var_det_is_heteroskedastic}%\\
%  avg \left[ \ln |C_v| \right] &= \ln |\Sigma_v| - d \cdot \ln{L}  + \sum^{d-1}_{i=0} \psi^0(L-i) \label{eqn:avg_log_det} \\
%  var \left[ \ln |C_v| \right] &=  \sum^{d-1}_{i=0} \psi^1(L-i) \label{eqn:var_log_det_is_homoskedastic}
\end{align}
}

@@

Similar to the way intensity-ratio is proposed as the discrimination measure for the multiplicative and heteroskedastic SAR intensity \citep{Rignot_1993_TGRS_896},
in this paper, the determinant-ratio and the change-ratio are also proposed as a discrimination measure for the POLSAR data.
  which is shown above to also suffer from the multiplicative and heteroskedastic phenomena.
%Similar to the way dispersion and contrast is defined for one-dimensional SAR \citep{Le_2010_ACRS},
%  this section introduces the consistent sense of distance for POLSAR.

@@

Statistical models have always been the foundation for deriving discrimination measures.
This is true for both fields of SAR and POLSAR.
For SAR, where the field is more or less settled on the issue,
  the statistical model for SAR intensity has been used to derive the most widely used intensity-ratio discrimination measure \citep{Rignot_1993_TGRS_896}.
For POLSAR, where the research field is much less mature,
  the same case should apply, 
  except that so far only asymptotic distributions have been derived for the most common foundation, i.e. the likelihood test statistics \citep{Conradsen_2003_TGRS_4}.

With the insight gained from section \ref{sec:sar_special_case_of_polsar}, this section presents a few results.
%  it should be clear that the proposed model can also be used to derive discrimination measures for POLSAR.
First, similar to the way that the statistical model for SAR intensity has been used to provide the foundation for SAR discrimination measures, e.g. intensity-ratio \citep{Rignot_1993_TGRS_896},
  the statistical model proposed for POLSAR covariance matrix determinant is to be reviewed as providing the foundation for POLSAR discrimination measures, i.e. the likelihood test statistics.
And second, new discrimination measures for POLSAR may be derived by learning from the existing SAR discrimination measures.

As for the first matter, in view of the models given in Eqn \ref{eqn:determinant_distribution},
  the likelihood test statistics presented in \citep{Conradsen_2003_TGRS_4} and rewritten in Eqns \ref{eqn:ori_likelyhood_stats} \& \ref{eqn:log_likelyhood_stats}
can be expressed as:
\begin{align*}
  \ln{Q} &\sim  k + L_x \Lambda^d_{L_x} + L_y \Lambda^d_{L_y} - (L_x + L_y) \Lambda^d_{(L_x + L_y)} \\
  Q &\sim e^k \frac{(\chi^d_{L_x})^{L_x} \cdot (\chi^d_{L_y})^{L_y}}{(\chi^d_{L_x + L_y})^{L_x + L_y}}   
\end{align*}
where $k = d \left[ (L_x + L_y) \ln(L_x + L_y) - L_x \ln{L_x} - L_y \ln{L_y} \right]$.
This, in essence, derives an exact statistical distribution for the likelihood test statistics,
  as opposed to the asymptotic distribution derived in \citep{{Conradsen_2003_TGRS_4}}.

@@

This section aims to verify the theoretical models given in Eqns. \ref{eqn:determinant_distribution}, \ref{eqn:determinant_ratio_distribution} and \ref{eqn:change_ratio_distribution} against practical data.
All of these models require the estimation of two parameters from the captured data.
These are (i) the dimensional number $d$ and (ii) the look number $L$.
The dimensional number $d$ is related to the type of (POL)SAR data captured,
  with $d=1,2,3$ corresponding to the cases of SAR, partial and full polarimetric SAR data, respectively.
The look-number $L$ can be taken either as nominally stated by the data provider,
  or it can be estimated using the technique proposed by \citep{Anfinsen_2009_TGRS_3795}.

@@

Since it has been shown in Section \ref{sec:sar_special_case_of_polsar} that the model for the case of $d=1$ matches exactly with the traditional model for SAR intensity,
  the validation of the proposed $d=1$ model is quite straightforward. %only a simple validation is presented for this case.  
Fig. \ref{fig:verify_POLSAR_model_1D} presents the results of a test performed for the stated purpose.
In the test, the intensity of single-channel SAR data (HH) for sample homogeneous areas in both datasets is extracted.
The histograms for both the intensity and intensity ratio %the log-distance and contrast
  are then plotted against the theoretical PDF given above.
%Also presented are the plots of the same test applied on a homogeneous patch inside the nine-look processed RADARSAT2 image.
The plots are obtained having set ENL to an estimated value obtained using the technique presented in \citep{Anfinsen_2009_TGRS_3795}.
In all cases, the good visual match between the actual data and model distribution tends to validate the proposed model. %providing us a simple validation of the result.

\begin{figure}[h]
%\captionsetup{justification=centering}
\centering
\begin{tabular}{c}
 \begin{minipage}[b]{2.8in} 
%	\subfloat[AIRSAR (HH) determinant]{
          \plotWithLegend{../images/verify_determinant_model_on_AIRSAR_1d.eps}{determinant}
%		 \epsfxsize=2.5in
%		 \epsfysize=2.5in
%		 \epsffile{../images/verify_determinant_model_on_AIRSAR_1d.eps}
		 \label{AIRSAR_1D_determinant}
                \subcaption{AIRSAR (HH) determinant}
 \end{minipage}                 
%	} 
	\hfill	
	%\subfloat[][RADARSAT2 (HH) \\ determinant]{
 \begin{minipage}[b]{2.8in} 
          \plotWithLegend{../images/verify_determinant_model_on_RADARSAT2_1d.eps}{determinant}
%		 \epsfxsize=2.5in
%		 \epsfysize=2.5in
%		 \epsffile{../images/verify_determinant_model_on_RADARSAT2_1d.eps} 	
		 \label{RADARSAT2_1D_determinant}
                \subcaption{RADARSAT2 (HH) determinant}
 \end{minipage} \\                
	%} \\
%	\subfloat[][AIRSAR (HH) \\ determinant ratio]{
 \begin{minipage}[b]{2.8in} 
          \plotWithLegend{../images/verify_det_ratio_model_on_AIRSAR_1d.eps}{determinant-ratio}
%		 \epsfxsize=1.5in
%		 \epsfysize=1.5in
%		 \epsffile{../images/verify_det_ratio_model_on_AIRSAR_1d.eps} 	
		 \label{AIRSAR_1D_det_ratio}
                \subcaption{AIRSAR (HH) determinant ratio}
 \end{minipage}                 
%	} 
	\hfill	
%	\subfloat[][RADARSAT2 (HH) \\ determinant ratio]{
 \begin{minipage}[b]{2.8in} 
          \plotWithLegend{../images/verify_det_ratio_model_on_RADARSAT2_1d.eps}{determinant-ratio}
%		 \epsfxsize=1.5in
%		 \epsfysize=1.5in
%		 \epsffile{../images/verify_det_ratio_model_on_RADARSAT2_1d.eps} 	
		 \label{RADARSAT2_1D_det_ratio}
                \subcaption{RADARSAT2 (HH) determinant ratio}
 \end{minipage} \\                 
%	} \\
%	\subfloat[][AIRSAR (HH) \\ change ratio]{
 \begin{minipage}[b]{2.8in} 
          \plotWithLegend{../images/verify_change_ratio_model_on_AIRSAR_1d.eps}{change-ratio}
%		 \epsfxsize=1.5in
%		 \epsfysize=1.5in
%		 \epsffile{../images/verify_change_ratio_model_on_AIRSAR_1d.eps} 	
		 \label{AIRSAR_1D_det_ratio}
                \subcaption{AIRSAR (HH) change ratio}
 \end{minipage}                 
%	}
	\hfill	
 \begin{minipage}[b]{2.8in} 
%	\subfloat[][RADARSAT2 (HH) \\ change ratio]{
%
\plotWithLegend{../images/verify_change_ratio_model_on_RADARSAT2_1d.eps}{change-ratio}
%		 \epsfxsize=1.5in
%		 \epsfysize=1.5in
%		 \epsffile{../images/verify_change_ratio_model_on_RADARSAT2_1d.eps} 	
%		 \label{RADARSAT2_1D_det_ratio}
                \subcaption{RADARSAT2 (HH) change ratio}
 \end{minipage}                 
%	}
\end{tabular}
\caption{The specific d=1 models are validated for both RADARSAT2 and AIRSAR datasets.
Each histogram plots the signal under investigation (along the x-axis) against its probabilistic distribution across the patch (on the y-axis).}
\label{fig:verify_POLSAR_model_1D}
\end{figure}
